---
title: "MLB"
author: "myTinyShinys"
output: 
  flexdashboard::flex_dashboard:
    
    theme: united
    orientation: rows
    social: [ menu ]
   # source_code: embed
runtime: shiny
---




```{r setup, include=FALSE}
# Load packages and initialize data here
library(flexdashboard)
library(shiny)
library(plotly)
library(readr)
library(lubridate)

library(XML)
library(stringr)



library(dplyr)

divRaces <- readRDS("allRes1969toDate.rds")
print(str(divRaces))
#divRaces <- read_feather("allRes1969toDate.feather") #issues with this. not suitable format for long term storage

Master <- read_csv("data/Master.csv")
Pitching <-  read_csv("data/Pitching.csv")

# create data for pitcher selectInput 
pitchers <- Master %>% 
  mutate(lastYear=ifelse(is.na(deathYear)," ",deathYear)) %>% 
          right_join(Pitching) %>% 
           select(nameFirst,nameLast,birthYear,lastYear,bbrefID,playerID) %>% 
           unique() %>% 
            arrange(bbrefID)

bbRef_pitchers <- pitchers$playerID

names(bbRef_pitchers) <- paste0(pitchers$nameFirst," ", pitchers$nameLast," (",pitchers$birthYear,"-",pitchers$lastYear,")")


## cf batters

  batters <- Master %>% 
    mutate(lastYear=ifelse(is.na(deathYear)," ",deathYear)) %>% 
    select(nameFirst,nameLast,birthYear,lastYear,bbrefID,playerID) %>% 
    unique() %>% 
    arrange(bbrefID)
  
  bbRef_batters <- batters$playerID
  
  names(bbRef_batters) <- paste0(batters$nameFirst," ", batters$nameLast," (",batters$birthYear,"-",batters$lastYear,")")
  

```

1994-2016 {data-navmenu="Division Races"}
===================================== 

Row {data-height=150}
-----------------------------------------------------------------------

```{r}
 inputPanel(sliderInput("race_yr","Enter year of choice",value=2016,min=1994,max=2016,sep=""),
            actionButton("go","Click for charts (takes 10+secs"))
```



Row
-----------------------------------------------------------------------

### AL East (Hover Lines for Details - Zoom as required)

```{r}
ALE <- eventReactive(input$go,{
 
  final_sel <- divRaces %>% 
    filter(league=="AL"&divison=="East"&season==input$race_yr) 
  
})

  
  output$ALE_race <- renderPlotly({
  plot_ly(ALE(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(ALE()$Team,
                       "<br>",ALE()$Date,
                       "<br> v ",ALE()$Opp," "," ",ALE()$R,"-",ALE()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",ALE()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("ALE_race")
```

### AL Central

```{r}
ALC <- eventReactive(input$go,{
  final_sel <- divRaces %>% 
    filter(league=="AL"&divison=="Central"&season==input$race_yr)
})

  
  output$ALC_race <- renderPlotly({
  plot_ly(ALC(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(ALC()$Team,
                       "<br>",ALC()$Date,
                       "<br> v ",ALC()$Opp," "," ",ALC()$R,"-",ALC()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",ALC()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("ALC_race")
   
```

### AL West

```{r}
ALW <- eventReactive(input$go,{
  final_sel <- divRaces %>% 
    filter(league=="AL"&divison=="West"&season==input$race_yr)
})

  
  output$ALW_race <- renderPlotly({
  plot_ly(ALW(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(ALW()$Team,
                       "<br>",ALW()$Date,
                       "<br> v ",ALW()$Opp," "," ",ALW()$R,"-",ALW()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",ALW()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("ALW_race")
```

Row
-----------------------------------------------------------------------

### NL East

```{r }
NLE <- eventReactive(input$go,{
  final_sel <- divRaces %>% 
    filter(league=="NL"&divison=="East"&season==input$race_yr)
})

  
  output$NLE_race <- renderPlotly({
  plot_ly(NLE(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(NLE()$Team,
                       "<br>",NLE()$Date,
                       "<br> v ",NLE()$Opp," "," ",NLE()$R,"-",NLE()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",NLE()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("NLE_race")
```

### NL Central

```{r}
NLC <- eventReactive(input$go,{
  final_sel <- divRaces %>% 
    filter(league=="NL"&divison=="Central"&season==input$race_yr)
})

  
  output$NLC_race <- renderPlotly({
  plot_ly(NLC(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(NLC()$Team,
                       "<br>",NLC()$Date,
                       "<br> v ",NLC()$Opp," "," ",NLC()$R,"-",NLC()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",NLC()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("NLC_race")
```

### NL West

```{r}
NLW <- eventReactive(input$go,{
  final_sel <- divRaces %>% 
    filter(league=="NL"&divison=="West"&season==input$race_yr)
})

  
  output$NLW_race <- renderPlotly({
  plot_ly(NLW(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(NLW()$Team,
                       "<br>",NLW()$Date,
                       "<br> v ",NLW()$Opp," "," ",NLW()$R,"-",NLW()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",NLW()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("NLW_race")
    
```

1969-1993 {data-navmenu="Division Races"}
===================================== 

Row {data-height=150}
-----------------------------------------------------------------------

```{r}
 inputPanel(sliderInput("race_yr2","Enter year of choice",value=1993,min=1969,max=1993,sep=""),
            actionButton("go2","Display Charts (coming soon)"))
```



Row
-----------------------------------------------------------------------

### AL East

```{r}
ALE2 <- eventReactive(input$go2,{
  final_sel <- divRaces %>% 
    filter(league=="AL"&divison=="East"&season==input$race_yr2)
})

  
  output$ALE2_race <- renderPlotly({
  plot_ly(ALE2(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(ALE2()$Team,
                       "<br>",ALE2()$Date,
                       "<br> v ",ALE2()$Opp," "," ",ALE2()$R,"-",ALE2()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",ALE2()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("ALE2_race")
```



### AL West

```{r}
ALW2 <- eventReactive(input$go2,{
  final_sel <- divRaces %>% 
    filter(league=="AL"&divison=="West"&season==input$race_yr2)
})

  
  output$ALW2_race <- renderPlotly({
  plot_ly(ALW2(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(ALW2()$Team,
                       "<br>",ALW2()$Date,
                       "<br> v ",ALW2()$Opp," "," ",ALW2()$R,"-",ALW2()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",ALW2()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("ALW2_race")
```

Row
-----------------------------------------------------------------------

### NL East

```{r }
NLE2 <- eventReactive(input$go2,{
  final_sel <- divRaces %>% 
    filter(league=="NL"&divison=="East"&season==input$race_yr2)
})

  
  output$NLE2_race <- renderPlotly({
  plot_ly(NLE2(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(NLE2()$Team,
                       "<br>",NLE2()$Date,
                       "<br> v ",NLE2()$Opp," "," ",NLE2()$R,"-",NLE2()$RA,
                       "<br> GB:",NLE2()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("NLE2_race")
```



### NL West

```{r}
NLW2 <- eventReactive(input$go2,{
  final_sel <- divRaces %>% 
    filter(league=="NL"&divison=="West"&season==input$race_yr2)
})

  
  output$NLW2_race <- renderPlotly({
  plot_ly(NLW2(), x = ~gameOrder, y = ~diff, mode = "lines", hoverinfo = "text",color=~Team,
          text = paste(NLW2()$Team,
                       "<br>",NLW2()$Date,
                       "<br> v ",NLW2()$Opp," "," ",NLW2()$R,"-",NLW2()$RA,
                      # "<br> W-L:",toDate,
                       "<br> GB:",NLW2()$GB)) %>% 
    
    layout(hovermode = "closest", 
           xaxis=list(title="Games Played"),
           yaxis=list(title="Games Behind Division Leader", autorange="reversed"
                  
           )
    ) %>% 
   config(displayModeBar = F,showLink = F)
  })
  
  plotlyOutput("NLW2_race")
    
```

Pitcher Game Scores
===================================== 
Inputs {.sidebar}
-------------------------------------

```{r}
includeMarkdown("gameScore.md")

inputPanel(
selectInput("log_Pitcher","Select or type Starter's Name",bbRef_pitchers, selected="hernafe02"),
actionButton("go3","Display Charts  (wait for it)")
)
includeMarkdown("gameScoreProviso.md")

```


```{r}
 

pgl <- eventReactive(input$go3,{
  
 print(input$log_Pitcher)
  print(glimpse(pitchers))
  
bbrefID <-pitchers %>% 
  filter(playerID==input$log_Pitcher) %>% 
  .$bbrefID

yrs <-Pitching %>% 
  filter(playerID==input$log_Pitcher)



career <- unique(yrs$yearID) # so combines teams

## takes account that lahman is only up to 2015
if(max(career)==2015) career=c(career,2016)


l <- length(career)
print(l)
x <- 0
withProgress(message=" ", value = 0, {
for (i in seq_along(career)) {

   print(career[i])
  incProgress(1/l, detail = paste("Season", career[i]))
print("here")
  
  url <- paste0("http://www.baseball-reference.com/players/gl.cgi?id=",bbrefID,"&t=p&year=",career[i])
 
  # url <- "http://www.baseball-reference.com/players/gl.cgi?id=priceda01&t=p&year=2016"
  tables <- readHTMLTable(url, stringsAsFactors=FALSE)
  tempdata <- tables[["pitching_gamelogs"]]
  
  ## when different number of columns appear can remove them
  
    if(!is.null(tempdata)) {
   x <- x+1
  if(career[i]>=2013) {
    print(career[i])
    tempdata <- tempdata %>% 
      select( -starts_with("DFS"))
  }
 
  
  tempdata$year <- career[i]
  
 
  if (x!=1) {
    data <- rbind(data,tempdata)
  } else {
    data <- tempdata
  }
}
}
})

# replace blank column heading
 names(data)[6] <- "venue"

data <-  data %>% 
        filter(Gcar!="Tm") %>% 
        filter(!is.na(Opp)) %>% 
        mutate(venue=ifelse(venue=="@",venue,"v")) %>% 
  mutate(Date=str_replace(Date,"Â","")) %>% 
  mutate(Gcar=as.integer(Gcar),GSc=as.integer(GSc))


return(data)

})

```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------


### Individual Game Scores <br> Zoom as required <br> Hover for details 

```{r}

output$pgl_individual <- renderPlotly({

 
  
  df <- pgl()
  

  
 df %>% 
  plot_ly() %>% 
    add_markers(x=~Gcar,y=~GSc,color=~Tm,
           hoverinfo = "text",
          #source = "gameScores",#not sure why that was in there
          text=~paste(Date," ",year,
                     "<br>",venue,Opp,
                     "<br>", Rslt,
                     "<br> IP:", IP,
                     "<br> Pitches:", Pit,
                     "<br> H:", H,
                     "<br> SO:", SO,
                     "<br> BB:", BB,
                     "<br> GS:", GSc)) %>%
    layout(hovermode = "closest",
           title=" ",
           xaxis=list(title="Career Game"),
           yaxis=list(title="Game Score",rangemode="tozero"
           )
    )
  })

plotlyOutput("pgl_individual")
```




### By Team

```{r}


output$pgl_team <- renderPlotly({
  pgl() %>% 
   plot_ly(y=~GSc,type="box",color=~Tm, boxpoints = "all", jitter = 0.3
           ) %>%
    layout(
           title=" ",
           xaxis=list(title="Team"),
           yaxis=list(title="Game Scores",rangemode="tozero"
           )
    )
  })

plotlyOutput("pgl_team")

```


### By Opposition

```{r}

output$pgl_opp <- renderPlotly({
  pgl() %>% 
    arrange(Opp) %>% 
  plot_ly(y=~GSc,type="box",color=~Opp, boxpoints = "all"
  ) %>%
  layout(
    title=" ",
    xaxis=list(title=" "),
    yaxis=list(title="Game Scores",rangemode="tozero"
    )
  )
  })

plotlyOutput("pgl_opp")

```

### By Year

```{r}

output$pgl_year <- renderPlotly({



  pgl() %>%
   plot_ly() %>% 
   add_boxplot(y=~GSc,color=~as.character(year), boxpoints = "all", jitter = 0.3
          ) %>%
    layout(showlegend=FALSE,
           title=" ",
           xaxis=list(title=" "),
           yaxis=list(title="Game Scores",rangemode="tozero"
           )
    )
  })

plotlyOutput("pgl_year")

```

### By Month

```{r}

output$pgl_month <- renderPlotly({


df <- pgl()

# some manipulation required - prob not most efficient!

 df$month <- tolower(str_sub(df$Date,1,3))
         df$day <- str_sub(df$Date,5,6)
         
         y <- paste0(df$day,df$month,df$year)
         
         df$trueDate<- as.Date(strptime(y, "%d%b%Y"))
         
         
         df$truemonth <- month(df$trueDate,label =TRUE, abbr=FALSE) # 
         
         df$numericmonth <- month(df$trueDate) # 
         
         df$Month <- factor(df$truemonth,levels=month.name)
         
       
         
       df  %>% 
           plot_ly() %>% 
         add_boxplot(y=~GSc,color=~Month, boxpoints = "all", jitter = 0.3) %>%
           layout(showlegend=FALSE,
                  title=" ",
                  xaxis=list(title=" ",autorange = "reversed"),
                  yaxis=list(title="Game Scores",rangemode="tozero"
                  )
           )
        
  })

plotlyOutput("pgl_month")

```

Pitcher WAR comparisons
===================================== 
Inputs {.sidebar}
-------------------------------------

```{r}
includeMarkdown("WAR.md")

inputPanel(
selectInput("war_Pitcher","Select or type Starter's Name",bbRef_pitchers, selected=c("kershcl01","martipe02"),multiple=T),
actionButton("go4","Display Charts  (wait for it)")
)


```


```{r}

pitcherWARdata <- eventReactive(input$go4,{
choices <- input$war_Pitcher

#withProgress(message=" ", value = 0, {
  for (i in seq_along(choices)) {
    
    print(choices[i])
    #incProgress(1/l, detail = paste("Pitcher", choices[i]))
    initial <- str_sub(choices[i],1,1)
    
    url <- paste0("http://www.baseball-reference.com/players/",initial,"/",choices[i],".shtml")
    
    # url <- "http://www.baseball-reference.com/players/gl.cgi?id=priceda01&t=p&year=2016"
    tables <- readHTMLTable(url, stringsAsFactors=FALSE)
    tempdata <- tables[["pitching_value"]]
    tempdata$bbrefID <- choices[i]
    
    
    
      
      if ( i!=1) {
        data <- rbind(data,tempdata)
      } else {
        data <- tempdata
      }
    }
#}

WARdata <-data %>% 
  left_join(Master) %>% 
  mutate(name=paste0(nameFirst," ",nameLast),WAR=as.numeric(WAR)) %>% 
  select(Age,name,WAR,bbrefID)

glimpse(WARdata)

# in caas 2 clubs
byYear <- WARdata %>% 
         group_by(Age,name,bbrefID) %>% 
         summarize(yrWAR=sum(WAR,na.rm=T)) %>% 
  group_by(name) %>% 
      mutate(cumWAR=cumsum(yrWAR))

info=list(WARdata=WARdata,byYear=byYear)
return(info)

})
```

Row
-----------------------------------------------------------------------

```{r pitcher WAR chart}

output$pitcherWAR <- renderPlotly({
  
  df <- pitcherWARdata()$byYear
   df  %>% 
     plot_ly() %>% 
  add_lines(x=~Age,y=~cumWAR,color=~name) %>% 
  add_bars(x=~Age,y=~yrWAR,color=~name,showlegend=FALSE) %>% 
  layout(
    title="Comparison of Annual and Cumulative Pitching bbRef WAR by Age",
   
    yaxis=list(title="Cumulative WAR")
  )
  
  
})

plotlyOutput("pitcherWAR")

```


Batter WAR comparisons
===================================== 
Inputs {.sidebar}
-------------------------------------

```{r}
includeMarkdown("WAR.md")

inputPanel(
selectInput("war_Batter","Select or type Batters' Name",bbRef_batters, selected=c("troutmi01","rodrial01"),multiple=T),
actionButton("go5","Display Charts  (wait for it)")
)


```


```{r}

 batterWARdata <- eventReactive(input$go5,{
    choices <- input$war_Batter
   
    
    #withProgress(message=" ", value = 0, {
    for (i in seq_along(choices)) {
      
      print(choices[i])
      #incProgress(1/l, detail = paste("Pitcher", choices[i]))
      initial <- str_sub(choices[i],1,1)
      
      url <- paste0("http://www.baseball-reference.com/players/",initial,"/",choices[i],".shtml")
      
      
      tables <- readHTMLTable(url, stringsAsFactors=FALSE)
      
      tempdata <- tables[["batting_value"]]
      tempdata$bbrefID <- choices[i]
      
      
      
      
      if ( i!=1) {
        data <- rbind(data,tempdata)
      } else {
        data <- tempdata
      }
    }
    #}
    
    WARdata <-data %>% 
      left_join(Master) %>% 
      mutate(name=paste0(nameFirst," ",nameLast),WAR=as.numeric(WAR)) %>% 
      select(Age,name,WAR,bbrefID)
    
    glimpse(WARdata)
    
    # in caas 2 clubs
    byYear <- WARdata %>% 
      group_by(Age,name,bbrefID) %>% 
      summarize(yrWAR=sum(WAR,na.rm=T)) %>% 
      group_by(name) %>% 
      mutate(cumWAR=cumsum(yrWAR))
    
    info=list(WARdata=WARdata,byYear=byYear)
    return(info)
})
```

Row
-----------------------------------------------------------------------

```{r batter WAR chart}

 output$batterWAR <- renderPlotly({
    
    df <- batterWARdata()$byYear #df <- byYear
    df  %>% 
      plot_ly() %>% 
      add_lines(x=~Age,y=~cumWAR,color=~name) %>% 
      add_bars(x=~Age,y=~yrWAR,color=~name,showlegend=FALSE) %>% 
      layout(
        title="Comparison of Annual and Cumulative Batting bbRef WAR by Age",
        
        yaxis=list(title="Cumulative WAR")
      )
    
    
  })
  
  plotlyOutput("batterWAR")

```

Batter gamelogs by Year
===================================== 
Inputs {.sidebar}
-------------------------------------

```{r gl1}
#includeMarkdown("WAR.md")

inputPanel(
selectInput("gl_batter","Select or type Batters' Name",bbRef_batters, selected=c("martiru01"))
#actionButton("go6","Display Charts  (wait for it)")
)


```

```{r gl1_yrs}

output$a <-renderUI({
  
  print("enter a")
  req(input$gl_batter)
  print(input$gl_batter)
  player <- input$gl_batter
  initial <- str_sub(player,1,1)
  
u1 <- paste0("http://www.baseball-reference.com/players/",initial,"/",player,"-bat.shtml")
print(u1)

  tables <- readHTMLTable(u1, stringsAsFactors=FALSE) #names(tables)
  print(names(tables))
  seasons <- tables[["batting_value"]]

yrs <-  seasons %>% 
    mutate(Year=as.integer(str_sub(Year,1,4))) %>% 
    arrange(desc(Year)) %>% 
    .$Year 
print(yrs)
  numericInput("gl_year","Select Year",value=max(yrs),min=min(yrs),max=max(yrs))

})

uiOutput("a")

```


Row
-----------------------------------------------------------------------



```{r gl_carpentry_plot}

output$gl <- renderPlotly({
  url <- paste0("http://www.baseball-reference.com/players/gl.cgi?id=",input$gl_batter,"&t=b&year=",input$gl_year)
  print(url)
  tables <- readHTMLTable(url, stringsAsFactors=FALSE)
  df <- tables[["batting_gamelogs"]]
 
  ## every variable has to be named
 colnames(df)[14] <- "double"
 colnames(df)[15] <- "triple"
 colnames(df)[6] <- "venue"
  
  
  df %>% 
    mutate(gameOrder=as.integer(Rk),OPS=as.numeric(OPS),H=as.integer(H),double=as.integer(double),triple=as.integer(triple),HR=as.integer(HR)) %>% 
    mutate(TB=4*HR+3*triple+2*double+H-HR-triple-double) %>% 
    mutate(bopper=ifelse(HR>0,"#ff0000","#fee3da")) %>% 
    plot_ly() %>% 
    add_lines(x=~gameOrder,y=~OPS, name='OPS',marker=list(color="ff0000", size=1)) %>% 
    add_bars(x=~gameOrder,y=~TB/50, name='Total Bases',color=~bopper,
             hoverinfo="text",
             text=~paste("TB:",TB,"<br>Opp:",Opp,"<br>",Date,"<br>Game",gameOrder)) %>% 
    layout(showlegend=FALSE)  %>% 
    config(displayModeBar = F,showLink = F)
})

plotlyOutput("gl")

```